---
title: "Week 12 - Recommendation Systems part 2"
author: "James Quacinella"
date: "11/10/2015"
output: pdf_document
---

# Exercise 9.4.2

If we wish to start out, as in Fig. 9.10, with all U and V entries set to the same value, what value minimizes the RMSE for the matrix M of our running example?

## Answer

We have a $5 x 5 M$ matrix which we want to decompose into $d=2$ dimensional $U$ and $V$
matricies.

"If we have chosen d as the lengths of the short sides of U and V , and a is the average nonblank element of M , then the elements of U and V should be sqrt(a/d)."

```{r}
d <- 2;
M <- matrix(c(5,2,4,4,3,3,1,2,4,1,2,NA,3,1,4,2,5,4,3,5,4,4,5,4,NA), ncol=5, byrow = TRUE);
sqrt(mean(M, na.rm = TRUE) / 2); # 1.276885
```

\pagebreak

# Exercise 9.4.3

```{r results='hold'}
# Starting off
U <- matrix(c(2.6, 1, 1.178, 1, 1, 1, 1, 1, 1, 1), ncol=2);
V <- matrix(c(1.617,1,1,1,1,1,1,1,1,1), ncol=5);

# Calculate P
P <- U %*% V;

# Show initital error
sqrt(sum((M - P)**2)); # 0.0002650962
```

## Answer

Starting with the U and V matrices in Fig. 9.16, do the following in order:

(a) Reconsider the value of $u_{11}$ . Find its new best value, given the changes that have been made so far.

```{r results='hold'}
max_j <- 5;
r <- 1;
s <- 1;

# Find the updated value by alculating the numerator and denominator
numerator <- sum(sapply(1:max_j, function(j) { V[s, j] * (M[r, j] - sum(U[r, -c(s)] * V[-c(s), j])); }))
denominator <- sum(sapply(1:max_j, function(j) { V[s, j]**2; }))
numerator
denominator

# Update and print U
U[r, s] <- numerator / denominator
U

# Show sum of squared error
sqrt(sum((M - U %*% V)**2)); # 0.0002333765

# Update P
P <- U %*% V;
```

(b) Then choose the best value for $u_{52}$

```{r results='hold'}
max_j <- 5;
r <- 5;
s <- 2;

# Find the updated value by alculating the numerator and denominator
numerator <- sum(sapply(1:max_j, function(j) { V[s, j] * (M[r, j] - sum(U[r, -c(s)] * V[-c(s), j])); }))
denominator <- sum(sapply(1:max_j, function(j) { V[s, j]**2; }))
numerator
denominator

# Update and print U
U[r, s] <- numerator / denominator
U

# Show sum of squared error
sqrt(sum((M - U %*% V)**2)); 

# Update P
P <- U %*% V;
```

(c) Then choose the best value for $v_{22}$.

```{r results='hold'}
max_i <- 5;
r <- 2;
s <- 2;

# Find the updated value by alculating the numerator and denominator
numerator <- sum(sapply(1:max_i, function(i) { U[i, r] * (M[i, s] - sum(U[i, -c(r)] * V[-c(r), s])); }))
denominator <- sum(sapply(1:max_i, function(i) { U[i, r]**2; }))
numerator
denominator

# Update and print U
V[r, s] <- numerator / denominator
V

# Show sum of squared error
sqrt(sum((M - U %*% V)**2));

# Update P
P <- U %*% V;
```
