---
title: "Week11 - Recoomendation Systems"
author: "James Quacinella"
date: "11/04/2015"
output: pdf_document
---

# Exercise 9.3.1

Figure 9.8 is a utility matrix, representing the ratings, on a 1–5 star scale, of eight items, a through h, by three users A, B, and C. Compute the following from the data of this matrix.

(a) Treating the utility matrix as boolean, compute the Jaccard distance between each pair of users.
(b) Repeat Part (a), but use the cosine distance.
(c) Treat ratings of 3, 4, and 5 as 1 and 1, 2, and blank as 0. Compute the Jaccard distance between each pair of users.
(d) Repeat Part (c), but use the cosine distance.
(e) Normalize the matrix by subtracting from each nonblank entry the average value for its user.
(f) Using the normalized matrix from Part (e), compute the cosine distance between each pair of users.

# Answer

Prep:

```{r}
require(lsa)

# Utility matrix for problem
utility.matrix <- data.frame(a=c(4, 0, 2), 
                            b=c(5, 3, 0),
                            c=c(0, 4, 1),
                            d=c(5, 3, 3),
                            e=c(1, 1, 0),
                            f=c(0, 2, 4),
                            g=c(3, 1, 5),
                            h=c(2, 0, 3))

utility.matrix <- matrix(c(4, 0, 2, 
                            5, 3, 0,
                            0, 4, 1,
                            5, 3, 3,
                            1, 1, 0,
                            0, 2, 4,
                            3, 1, 5,
                            2, 0, 3), nrow=3)


#  Convert user row into boolean vector
make.boolean <- function(user) {
  return(lapply(user, function(x) {return(as.integer(x != 0));}));
}

# Jaccard distance
dist.jaccard <- function(x, y) {
  return(1 - sum(x & y) / sum(x | y));
}

# Cosine distance
dist.cosine <- function(x, y) {
  return(1 - cosine(x,y));
}


# For part a), create boolean version of utility matrix
utility.matrix.boolean <- data.frame()
utility.matrix.boolean <- rbind(utility.matrix.boolean, make.boolean(utility.matrix[1,]))
utility.matrix.boolean <- rbind(utility.matrix.boolean, make.boolean(utility.matrix[2,]))
utility.matrix.boolean <- rbind(utility.matrix.boolean, make.boolean(utility.matrix[3,]))
colnames(utility.matrix.boolean) <- c('a', 'b', 'c', 'd', 'e', 'f', 'g', 'h')

# Example of Jaccard distances working (from book)
#dist.jaccard(c(4,0,0,5,1,0,0), c(5,5,4,0,0,0,0)) # 0.8
#dist.jaccard(c(4,0,0,5,1,0,0), c(0,0,0,2,4,5,0)) # 0.5
```

a) Treating the utility matrix as boolean, compute the Jaccard distance between each pair of users

```{r}
dist.jaccard(utility.matrix.boolean[1,], utility.matrix.boolean[2,])
dist.jaccard(utility.matrix.boolean[1,], utility.matrix.boolean[3,])
dist.jaccard(utility.matrix.boolean[2,], utility.matrix.boolean[3,])
```

b) Repeat Part (a), but use the cosine distance.

```{r}
# cosine(c(4,0,0,5,1,0,0), c(5,5,4,0,0,0,0)) # 0.37986
# cosine(c(4,0,0,5,1,0,0), c(0,0,0,2,4,5,0)) # 0.322

dist.cosine(as.matrix(utility.matrix)[1,], as.matrix(utility.matrix)[2,]) # 0.6010408
dist.cosine(as.matrix(utility.matrix)[1,], as.matrix(utility.matrix)[3,]) # 0.6149187
dist.cosine(as.matrix(utility.matrix)[2,], as.matrix(utility.matrix)[3,]) # 0.5138701
```

c) Treat ratings of 3, 4, and 5 as 1 and 1, 2, and blank as 0. Compute the Jaccard distance between each pair of users.


```{r}
# Treat ratings of 3, 4, and 5 as 1 and 1, 2, and blank as 0.
make.grouped <- function(user) { 
  return(lapply(user, function(x) {
    if(x==0||x==1||x==2) {
      return(0);
    } else {
      return(1);
    }
  }));
}

utility.matrix.grouped <- data.frame()
utility.matrix.grouped <- rbind(utility.matrix.grouped, make.grouped(utility.matrix[1,]))
utility.matrix.grouped <- rbind(utility.matrix.grouped, make.grouped(utility.matrix[2,]))
utility.matrix.grouped <- rbind(utility.matrix.grouped, make.grouped(utility.matrix[3,]))
colnames(utility.matrix.grouped) <- c('a', 'b', 'c', 'd', 'e', 'f', 'g', 'h')

utility.matrix.grouped

dist.jaccard(utility.matrix.grouped[1,], utility.matrix.grouped[2,]) # 0.6
dist.jaccard(utility.matrix.grouped[1,], utility.matrix.grouped[3,]) # 0.6666667
dist.jaccard(utility.matrix.grouped[2,], utility.matrix.grouped[3,]) # 0.8333333
```

d) Repeat Part (c), but use the cosine distance.

```{r}
dist.cosine(as.matrix(utility.matrix.grouped)[1,], as.matrix(utility.matrix.grouped)[2,]) # 0.5773503
dist.cosine(as.matrix(utility.matrix.grouped)[1,], as.matrix(utility.matrix.grouped)[3,]) # 0.5
dist.cosine(as.matrix(utility.matrix.grouped)[2,], as.matrix(utility.matrix.grouped)[3,]) # 0.2886751
```

e) Normalize the matrix by subtracting from each nonblank entry the average value for its user.

```{r}
utility.matrix.normalized <- utility.matrix;
utility.matrix.normalized[1,] <- utility.matrix.normalized[1,] - mean(utility.matrix.normalized[1,]);
utility.matrix.normalized[2,] <- utility.matrix.normalized[2,] - mean(utility.matrix.normalized[2,]);
utility.matrix.normalized[3,] <- utility.matrix.normalized[3,] - mean(utility.matrix.normalized[3,]);
utility.matrix.normalized
```

f) Using the normalized matrix from Part (e), compute the cosine distance between each pair of users.

```{r}
dist.cosine(as.matrix(utility.matrix.normalized)[1,], as.matrix(utility.matrix.normalized)[2,]) # 1.046374
dist.cosine(as.matrix(utility.matrix.normalized)[1,], as.matrix(utility.matrix.normalized)[3,]) # 1.037662
dist.cosine(as.matrix(utility.matrix.normalized)[2,], as.matrix(utility.matrix.normalized)[3,]) # 1.288179
```