---
title: "Week 13 - Social Network Mining"
author: "James Quacinella"
date: "11/14/2015"
output: pdf_document
---



# 10.2.1

Figure 10.9 is an example of a social-network graph. Use the Girvan-Newman approach to find the number of shortest paths from each of the following nodes that pass through each of the edges. (a) A (b) B.

## Answer 

(a)

```{r fig.width=7, fig.height=10,echo=FALSE}
library(jpeg)
library(grid)
img <- readJPEG("/home/james/Development/Masters/IS622/Week13/pica.jpg")
grid.raster(img)
```

(B) 

```{r fig.width=7, fig.height=10,echo=FALSE}
img <- readJPEG("/home/james/Development/Masters/IS622/Week13/picb.jpg")
grid.raster(img)
```

\pagebreak 

# 10.2.2 (section 10.2.6)

Using symmetry, the calculations of Exercise 10.2.1 are all you need to compute the betweenness of each edge. Do the calculation.

## Answer

Having the diagram start with node A is the same diagram that would result if the root was I or F. That means that diagram can be counted on 3 times in the summation calculation. The other diagram with root node B is the same that would result for all the other 5 nodes for a total of 6 (B,C,D,E,H,I). Here is a listing of all the edges and thir betweeness measurements:

A - B: 
  - When A is root: 4
  
  - When F is root: 0
  
  - When I is root: 1
  
  - When B is root: 1 
  
  - When C is root: 
  
  - When D is root: 
  
  - When H is root: 
  
  - When G is root: 
  
  - When E is root: 
  

A - C: Same as A-B due to symmetry

B - C:

B - H: 

C - D: 

H - I: Same as A-B due to symmetry

H - G: 

I - G: Same as A-B due to symmetry

G - E: 

D - E: 

D - F: Same as A-B due to symmetry

E - F: Same as A-B due to symmetry

\pagebreak 

# Exercise 10.4.1

For the graph of Fig. 10.9, construct:

(a) The adjacency matrix.
(b) The degree matrix.
(c) The Laplacian matrix.

## Answer

```{r}
# Setup the degree matrix
D = matrix(c(2,0,0,0,0,0,0,0,0,
             0,3,0,0,0,0,0,0,0,
             0,0,3,0,0,0,0,0,0,
             0,0,0,3,0,0,0,0,0,
             0,0,0,0,3,0,0,0,0,
             0,0,0,0,0,2,0,0,0,
             0,0,0,0,0,0,3,0,0,
             0,0,0,0,0,0,0,3,0,
             0,0,0,0,0,0,0,0,2), ncol=9)

# Setup col and row names
colnames(D)<-c('A','B','C','D','E','F','G','H','I')
rownames(D)<-c('A','B','C','D','E','F','G','H','I')

# Write up the adjanceny matrix
          #  A B C D E F G H I
A = matrix(c(0,1,1,0,0,0,0,0,0, # A
             1,0,1,0,0,0,0,1,0, # B
             1,1,0,1,0,0,0,0,0, # C
             0,0,1,0,1,1,0,0,0, # D
             0,0,0,1,0,1,1,0,0, # E
             0,0,0,1,1,0,0,0,0, # F
             0,0,0,0,1,0,0,1,1, # G
             0,1,0,0,0,0,1,0,1, # H
             0,0,0,0,0,0,1,1,0), ncol=9)

# Setup col and row names
colnames(A)<-c('A','B','C','D','E','F','G','H','I')
rownames(A)<-c('A','B','C','D','E','F','G','H','I')

# Calculate Laplacian
L = D - A
```

\pagebreak

```{r}
# Print results
D
A
L
```