---
title: "Week 4"
author: "James Quacinella"
date: "09/27/2015"
output: pdf_document
---

Exercise 3.1.3 : Suppose we have a universal set U of n elements, and we
choose two subsets S and T at random, each with m of the n elements. What
is the expected value of the Jaccard similarity of S and T ?

\pagebreak

# Exercise 3.3.3

In Fig. 3.5 is a matrix with six rows.

$$\begin{array}
{rrrrr}
element & S_1 & S_2 & S_3 & S_4 \\
0 & 0 & 1 & 0 & 1 \\
1 & 0 & 1 & 0 & 0 \\
2 & 1 & 0 & 0 & 1 \\
3 & 0 & 0 & 1 & 0 \\
4 & 0 & 0 & 1 & 1 \\
5 & 1 & 0 & 0 & 0
\end{array}
$$


(a) Compute the minhash signature for each column if we use the following
three hash functions: $h_1(x) = 2x + 1 mod 6; h_2(x) = 3x + 2 mod 6; h_3(x) = 5x + 2 mod 6.$

(b) Which of these hash functions are true permutations?

(c) How close are the estimated Jaccard similarities for the six pairs of columns
to the true Jaccard similarities?


## Answer

(a) Lets create a matrix with the has function computed, just ilke page 82 in the textbook:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 & h_1(x) & h_2(x) & h_3(x) \\
0 & 0 & 1 & 0 & 1 & 1 & 2 & 2 \\
1 & 0 & 1 & 0 & 0 & 3 & 5 & 1 \\
2 & 1 & 0 & 0 & 1 & 5 & 2 & 0 \\
3 & 0 & 0 & 1 & 0 & 1 & 5 & 5 \\
4 & 0 & 0 & 1 & 1 & 3 & 2 & 4 \\
5 & 1 & 0 & 0 & 0 & 5 & 5 & 3
\end{array}
$$

So far so good. Next we initialize a matrix that consists of all $\inf$:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & \infty & \infty & \infty & \infty \\
h_2(x) & \infty & \infty & \infty & \infty \\
h_3(x) & \infty & \infty & \infty & \infty
\end{array}
$$

Looking at row 0, we have 1's for $S_2$ and $S_4$ so only those columns can change. Since the values from the hash colums are all smaller than $\infty$, the matrix looks like this:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & \infty & 1 & \infty & 1 \\
h_2(x) & \infty & 2 & \infty & 2 \\
h_3(x) & \infty & 2 & \infty & 2
\end{array}
$$

Next row, we follow the same procedure. In this case, we have 1's for only $S_2$, and only the value for $h_3(x)$ changes, since its the only value less than what we have in the matrix currently:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & \infty & 1 & \infty & 1 \\
h_2(x) & \infty & 2 & \infty & 2 \\
h_3(x) & \infty & 1 & \infty & 2
\end{array}
$$

For row with label 2, we get:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & 5 & 1 & \infty & 1 \\
h_2(x) & 2 & 2 & \infty & 2 \\
h_3(x) & 0 & 1 & \infty & 0
\end{array}
$$

For row with label 3, we get:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & 5 & 1 & 1 & 1 \\
h_2(x) & 2 & 2 & 5 & 2 \\
h_3(x) & 0 & 1 & 5 & 0
\end{array}
$$

For row with label 4, we get:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & 5 & 1 & 1 & 1 \\
h_2(x) & 2 & 2 & 2 & 2 \\
h_3(x) & 0 & 1 & 4 & 0
\end{array}
$$

For the last row, our final min-hash signature matrix looks like:

$$\begin{array}
{rrrrrrrr}
row & S_1 & S_2 & S_3 & S_4 \\
h_1(x) & 5 & 1 & 1 & 1 \\
h_2(x) & 2 & 2 & 2 & 2 \\
h_3(x) & 0 & 1 & 4 & 0
\end{array}
$$


(b) Only $h_3(x)$ is a hash function that is a true permutations since there are no collisions amoung the outputs for any of the rows. The other functions have duplicate output values.

(c) Lets calculate the real Jaccard similarities, which is the intersection carindality divided by the union cardinality:

$$\begin{array}
{cc}
SIM_{real}(S_1, S_2) = \frac{0}{4} & SIM_{approx}(S_1, S_2) = \frac{1}{3} \\
SIM_{real}(S_1, S_3) = \frac{0}{4} & SIM_{approx}(S_1, S_3) = \frac{1}{3} \\
SIM_{real}(S_1, S_4) = \frac{1}{4} & SIM_{approx}(S_1, S_4) = \frac{2}{3} \\
SIM_{real}(S_2, S_3) = \frac{0}{4} & SIM_{approx}(S_2, S_3) = \frac{2}{3} \\
SIM_{real}(S_2, S_4) = \frac{1}{4} & SIM_{approx}(S_2, S_4) = \frac{2}{3} \\
SIM_{real}(S_3, S_4) = \frac{1}{4} & SIM_{approx}(S_3, S_4) = \frac{2}{3}
\end{array}
$$

\pagebreak

# Exercise 3.5.5

Compute the cosines of the angles between each of the following pairs of vectors:

(a) (3, -1, 2) and (-2, 3, 1).
(b) (1, 2, 3) and (2, 4, 6).
(c) (5, 0, -4) and (-1, -6, 2).
(d) (0, 1, 1, 0, 1, 1) and (0, 0, 1, 0, 0, 0).

## Answer

The cosine distance metric is the vector dot product divided by the product of the vector $L_2$-norms:

(a) $x = (3, -1, 2) and y = (-2, 3, 1)$.
The $L_2$ norm for $x$ is $\sqrt(3^2 + (-1)^2 + 2^2) = \sqrt(14)$

The $L_2$ norm for $y$ is $\sqrt((-2)^2 + 3^2 + 1^2) = \sqrt(14)$

The dot product $x \dot y$ is $(3)(-2) + (-1)(3) + (2)(1) = - 6 - 3 + 2 = -7$

Therefore the cosine of the angle between $x$ and $y$ is $\frac{-7}{\sqrt(14)\sqrt(14)} = -0.5$. 


(b) $x = (1, 2, 3) and y = (2, 4, 6)$.
The $L_2$ norm for $x$ is $\sqrt(1^2 + 2^2 + 3^2) = \sqrt(14)$

The $L_2$ norm for $y$ is $\sqrt(2^2 + 4^2 + 6^2) = \sqrt(56)$

The dot product $x \dot y$ is $(1)(2) + (2)(4) + (3)(6) = 28$

Therefore the cosine of the angle between $x$ and $y$ is $\frac{28}{\sqrt(14)\sqrt(56)} = 1$.


(c) $x = (5, 0, -4) and y = (-1, -6, 2).$.
The $L_2$ norm for $x$ is $\sqrt(5^2 + 0^2 + (-4)^2) = \sqrt(41)$

The $L_2$ norm for $y$ is $\sqrt((-1)^2 + (-6)^2 + 2^2) = \sqrt(41)$

The dot product $x \dot y$ is $(5)(-1) + (0)(-6) + (-4)(2) = -13$

Therefore the cosine of the angle between $x$ and $y$ is $\frac{-13}{\sqrt(41)\sqrt(41)} = -0.317$.


(d) $x = (0, 1, 1, 0, 1, 1) and y = (0, 0, 1, 0, 0, 0).$.
The $L_2$ norm for $x$ is $\sqrt(4) = 2$

The $L_2$ norm for $y$ is $\sqrt(1)$

The dot product $x \dot y$ is $(0)(0) + (1)(0) + (1)(1) + (0)(0) + (1)(0) + (1)(0) = 1$

Therefore the cosine of the angle between $x$ and $y$ is $\frac{1}{2}$.



\pagebreak

# Exercise 3.7.1

Suppose we construct the basic family of six locality-sensitive
functions for vectors of length six. For each pair of the vectors 000000, 110011,
010101, and 011100, which of the six functions makes them candidates?

## Answer

Lets define the 6 hash functions as $h_i(x)$, which returns the $i$-th bit in the vector $x$. Therefore, two values will collide, or make them candidates for similarity testing, when they agree on their $i$-th value.

So, the pairs are as follows:

000000 and 110011 will have $h_3$ and $h_4$ make them candidates

000000 and 010101 will have $h_1$, $h_3$ and $h_5$ make them candidates

000000 and 011100 will have $h_1$, $h_5$ and $h_6$ make them candidates

110011 and 010101 will have $h_2$, $h_3$ and $h_6$ make them candidates

110011 and 011100 will have $h_2$ make them candidates

010101 and 011100 will have $h_1$, $h_2$, $h_4$ and $h_5$ make them candidates

\pagebreak

Code: Write an R function `shingle(x, k)` that generates k-shingles from a given character vector x.


```{r}
require("sets");

# Terrible implementation of shingles() function
## Loops over starting positions and size k, generating substrings
## and returning a set of k-shingles
shingles <- function(x, k) {
  # Create empty set
  s <- canonicalize_set_and_mapping(c());
  
  # Loop over every position in string ...
  for (i in 1:nchar(x)) {
    # Loop over various shingle sizes ...
    for (j in 1:k) {
      # Then create shingles starting from position with size j
      # and do set-union with set
      s <- s + set(substring(x, i, i+j-1));
    }
  }
  
  # Return set
  s
}
```

```{r}
# Slightly better but could be simpler without use of unlist, etc
shingles2 <- function(x, k) {
  canonicalize_set_and_mapping(unlist(lapply(1:nchar(x), function(l) { lapply(1:k, function(k_shin) {as.character(substr(x, l, l+k_shin))})})))
}
```

Examples:

```{r}
print(shingles("james", 2));
print(shingles("This is the title of some article", 5));
```

```{r}
print(shingles2("james", 2)$set);
print(shingles2("This is the title of some article", 5)$set);
```